### 2015-04-08 Final_Paper/Paper.tex
This is the actual draft of the Introduction. =D

\documentclass[12pt]{article}
\usepackage{color}
\begin{document}
\begin{titlepage}
\title{\textbf{CPSC 504 Project\\\vspace{3 cm}\Huge{An investigation into an application to manage data transformations and updates}\vspace{2 cm}}}
\date{\vspace{2 cm} April 14, 2015}

\author{
 \makebox[1.0\linewidth]{Laura Cang}\\cang@cs.ubc.ca\\
  \and \makebox[1.0\linewidth]{Kailang Jiang}\\jiangkl@cs.ubc.ca\\
  \and \makebox[1.0\linewidth]{Jessica Wong}\\jhmwong@cs.ubc.ca\\ 
}
\maketitle
\thispagestyle{empty}
\end{titlepage}

\newpage
\section{Introduction}
This project is built on top of the theoretical framework done in the M.Sc thesis of Arni Thrastarson \cite{arniThesis} and focuses on the first of the four scenarios detailed in the work. The first scenario is about trying to investigate how changes made by a data application can be captured and propagated to a remote data source. The main workflow of this scenario would be a remote (possibly third party) application pulling the data from a data source to use locally. The application will receive a local copy of the data along with a schema mapping to identify which tables the columns in the local copy of the data came from. A user using the application may notice inconsistencies with the data and go to correct it. The changes made by this user are logged in a transformation script and when the time comes to update the data source, the transformation script is sent to the data store and the changes detailed in the script are applied.

\noindent Our primary contributions are:
\begin{enumerate}
	\item{Using logical programming concepts to define a data transformation language that captures changes made to a relational data model}
	\item{Design decisions about how the relational model should react and handle schema changes}
	\item{A preliminary investigation into view maintenance policies for the remote datastore}
	\item{Further detailing the scenario in which an application views a subset of columns from one or more tables in a remote dataset}
	\item{Deciding which issues to assign to the server side and which to the client side}
\end{enumerate}

\section{Motivation}
In our project, our main focus is on a scenario where the user views any subset of columns from one or more tables, it is important to figure out how this relates to the real world.

\subsection{Example Scenario}
In our project, we decided to borrow the dataset used in CPSC 304, an introductory relational database course at UBC, to use as the test dataset in our application. This was a movie dataset that we decided to take three columns out of to use

\section{Solution}
In this project, we focus on three different ways a user can use the application to fetch data.

\noindent

\section{Types of Changes Supported}
\subsection{Transformation Language}
\subsection{View Maintenance}
\subsection{Proof of Concept}
In the proof of concept, we had an extra step that was not included in \cite{arniThesis}. We decided to allow for the user to choose which changes to sync to the remote data source. \textcolor{red}{add screen shots!! Kailang can write about all the cool things she did here!! :)}

\section{Related Work}

\section{Future Work}
Future extensions of this work can look into testing the transformation language on other data formats like XML to see if it can effectively capture the changes that happen in an XML document. We can also look into how to handle situations where there are multiple concurrent users drawing upon the same data source and what policies should be adopted into order to achieve the best balance between performance and data consistency. Updating the data source too often could cause performance overhead from locking up the data source while updating too infrequently could cause inconsistent data to be propagated to many different users. The amount of effort and operational overhead involved with many people making the same changes would be undesirable.

Another extension of the project could be investigating how the data transformations could be integrated into data provenance to help with data conflict resolutions \cite{arniThesis}. As the project currently stands, the remote application can push any type of data to the data source. If the data being pushed to the data source violates any table constraints (e.g., if a user tries to push a non-unique value into a column that has specified that it only takes unique values), the database administrator has to manually resolve that violation. It is possible to investigate whether or not provenance information can be leveraged to help make these decisions thus lowering the workload on the database administrator. A possible example of how to use the provenance information could be using the previous history of the sorts of changes have been accepted for that column or table to determine the likelihood of the current data violation being accepted. Provenance could also be used to determine the likelihood of correctness by the app or user who made the change; as information about who has changed values can be tracked, it is not inconceivable that the statistics of who has made the best or the most correct/acceptable changes can be used in some way.

\bibliographystyle{abbrv}
\bibliography{final_paper_biblography}

\end{document}